DTOADQ

### 2017-11-29 testdocs/test.tex
doicumentation

  The rendering equation is:
  \begin{align}
    L_o(P, \omega_o) = L_e(P, \omega_o) + \int_{S}f_s(P, \omega_i,
    \omega_o) L_i(P, \omega_i) \vert cos\theta_i\vert d\omega_i
  \end{align}

  Where:
  	 $Lo(P,\:\omega_o)$ is the radiance at point $P$ in space towards direction $\omega_o$.

  	 $L_e(P,\:\omega_o)$ is the emittance at point $P$ towards $\omega_o$ (such as a lamp).

  	 $S$ is the unit sphere in which the angle $\omega_i$ is integrated over in respects to the normal at $P$. Also, this means, implicitly, that the equation is in respects to solid angle; the angle at which a primitive $A$ subtends $B$.

  	 $f_s$ is the Bidirectional Scattering Distribution Function. This will be detailed in the BSDF chapter, suffice it to say, it describes how photons are scattered (or absorbed) from point P in space from direction $\omega_i$ towards $\omega_o$ -- either through reflection (BRDF), or transmittance (BTDF). In volumetric path tracing, there is also subsurface scattering.

  	 $L_i(P, \omega_i)$ is the rendering equation at P from $\omega_i$

  	 And $cos\theta_i$ is the normal attenuation.

  Conceptually, this is the behavior of photons bouncing off several surfaces of a scene. While it's a very simple model, it's impossible to solve analytically for anything but the most simple of scenes, mostly due to, for example, the radiance at $P0$ relies on $P1$, and $P1$ relies on $P0$. Another thing to remember is that while it is shown to be integrated over a sphere, it is usually shown to be integrated over hemisphere $\Omega$ using either $f_r$ or $f_s$.

@@

  \begin{list}{}{}
  \item BSDF Spectrum: The albedo at $(P,\:\omega_i,\:\omega_o)$
  \item BSDF Sample: Samples a $\omega_o$ for a $(P,\:\omega_i)$
  \item BSDF PDF: The probability distribution of $(P,\:\omega_i,\:\omega_o)$
  \end{list}

  There are many more optimizations to be made as well. For example, at each scattering event, an additional photon can be sent towards each emitter. This is what is known as next-event estimation. While this improves performance drastically, naively averaging each contribution is a problem, because each path has a different probability to their paths/contributions. The technique to sample them properly, multiple importance sampling, is related to BDPT, however it will be detailed later. The important thing to understand is that, in this sampling strategy, there is a new PDF introduced which relates to the probability of a path being generated, and this is critical to understand the concept of BDPT.

  Now, the optimization that is of focus in DTQ is Bidirectional Path Tracing. Instead of starting from an eye or light, a 'photon' (path) is generated from both, allowed to scatter in the scene for a few paths (the length is determined by Russian Roulette described later), and connected. It's not exactly the end-points of the paths that are being connected, but each vertex generated is connected to each other. [SHOW EXAM{E}

  ]This actually has an interesting correlation to next-event estimation, as it could be seen as 'extending' the amount of emitters that can be connected to.

  Unfortunately, this formula does not work with BDPT, as it is necessary to transform the formula to work in terms of surface areas rather than solid angle, ei, $v_0 \rightarrow v_1$. The transformation looks like so:
  \begin{align}
    f_s(v \rightarrow \dot{v} \rightarrow \ddot{v}) &= f_s(P, \omega_i,
                                                      \omega_o)\\
    L_e(v \rightarrow \dot{v}) &= L_e(P, \omega_o)\\
    L_e(v \rightarrow \dot{v}) &= L_e^0(v) L_e^1(v \rightarrow \dot{v})
  \end{align}
    That is, instead of looking at a point $P$ with an incoming and outgoing
    angle $(\omega_i,\:\omega_o)$, look at the path from $P_{-1}$ ($v$) to
    $P_{+1}$ ($\ddot{v}$). More accurately, it's the transport between
    three different surfaces of the scene. There's a small change to be made in the case of $L_e$, as DTOADQ [at least, for now], only has support for area-light. Thus consider:

@@

    Instead of working with solid angles, we now work with a geometry term, G, that converts a PDF with respect to solid angle to surface area, which involves the inverse squared distance, and the cosine angle between the geometric normals at $v$ and $\dot{v}$:

@@

    Where $V$ is the visibility test. Thus the rendering equation with respects
    to area is:
  \begin{align}
    L(v \rightarrow \dot{v}) = L_e(v \rightarrow \dot{v}) \int_ML(v \rightarrow
    \dot{v}) G(v\leftrightarrow \dot{v}) f_s(v\rightarrow \dot{v} \rightarrow
    \ddot{v}) d_A
  \end{align}
    where $A$ is the area on $M$, and $M$ is the union of all scene surfaces.
    This is known as the three-point form or the light transport equation. By
    recursively expanding$^{[1]}$, so that the three point form is integrated
    over every possible set of possible paths, this can be rewritten as the more
    usable format:
  \begin{align} I_j = \int_\Omega f_j(\bar{v}) d\mu(\bar{v}) \end{align}
    Where $\bar{v}$ is a path $z_0\rightarrow z_s \rightarrow y_t \rightarrow
    y_0 (eye \rightarrow light)$, $\Omega$ is the combination of all possible
    paths of any length, and $\mu$ is the area-product measure [the product of
    all the expanded $d_A$]. Special care has to be taken with the definition of
    path itself, as in the source code, there never exists a path $\bar{v}$,
    only eye-path $\bar{z}$ and light-path . Right now, this is analytically
    unsolvable. To compute this, we need to apply monte carlo sampling and limit
    the maximum path length:
  \begin{align}
    I_j \approx \frac{1.0f}{N}\Sigma_0^NF\\
    F = \Sigma_s\Sigma_t\frac{f_j(\bar{v}_{s, t})}{p(\bar{v}_{s, t})}
  \end{align}

@@

    And now
  \begin{align}
    F = \Sigma_s\Sigma_t \alpha^L_s c_{s, t} \alpha^E_t
  \end{align}
  \begin{align}
    \alpha^{L|E}_i =
    \begin{cases}
      \frac{L_e^0(y_0 \rightarrow y_1)}{p_A(y_1)} | \frac{W^0_e(z_0)}{p_A(z_0)},
      &i = 1,\\
      (\frac{f_s(y_{i+1} \rightarrow y_i \rightarrow y_{i-1})}{p_{\sigma}(y_i
      \rightarrow y_{i-1})} | \frac{f_s(z_{i-1} \rightarrow z_{i-2} \rightarrow
      z_{i-3}}{p_\sigma(z_{i-2} \rightarrow z_{i-1})}) \alpha^{L|E}_{i-1},
      &i > 1
    \end{cases}
  \end{align}
    There is no case for $i = 0$ as either, the path doesn't exist, or if one
    does exist, $y_0$ and $z_0$ do not contribute to the image. Veach (and many other
    sources), leave this in, but it's unnecessarily confusing. A special case
    needs to be handled, unfortunately, for $\alpha^E_0$ as it is being
    generated on the fly while $\bar{z}$ is being constructed. The two options
    is to unroll the first iteration of the construction loop, or just allow the
    special case to exist. The latter is the better choice, as all kernels will
    enter and exit the special case at the same time, and the GPU might just
    unroll the loop anyways.
    There may be cases where $s = 0$ ($t = 0$ is not possible without a
    physical camera lens), which is equivalent to just forward path tracing.
    There is also to consider, what happens on $s = 1$ and $t = 1$, %TODO
    Well, in either case where both paths exist or only the camera path exists,
    the connection strategy below describes how to connect the edges of the two
    paths
  \begin{align}
    c_{s, t} =
    \begin{cases}
      L_e(z_{-2} \rightarrow z_{-1}), &s = 0, t > 1,\\
      f_s(y_{-2} \rightarrow y_{-1} \rightarrow z_{-1}) G(y_{-1} \leftrightarrow
      z_{-1}) f_s(y_{-1} \rightarrow z_{-1} \rightarrow z_{-2}), &s > 0, t > 1
    \end{cases}
  \end{align}
    Is the $s = 0, t > 1$
    strategy worthwhile to compute? Those familiar with next event
    estimation would know that there are situations in which this is beneficial.
    Take, for example, a glossy surface with a small solid angle subtending to
    the surface of a large light source. The $s=1$ sample strategy in this case
    would only contribute when the $L0$ vertex is within that solid angle.
    The effective strategy to sample in this case is $s = 0$.
    In code, $\alpha_i^E\:c_{s,t}\:\alpha_0^L$ looks like (MIS not shown):
    \begin{lstlisting}
     float3 wi = normalize(E1.origin - E0.origin),
            wo = normalize(L1.origin - E1.origin);
     float3 albedo = _BRDF_F(wi, E1.normal, wo, E1.material, E1.mat_col);
     albedo /= BSDF_PDF(wo, wi, &E1);
     Spectrum contribution = E1.radiance * albedo;
     contribution *= Geometric_Term(&E1, &L1, si, Tx);
     contribution *= tlight.emission/Light_PDF(&tlight);
    \end{lstlisting}

@@

  The Geometric term is used to control how microfacets 'shadow' each other. The DTQ BRDF uses Heits' [2014] SmithGGXCorrelated modified to use a half vector, combined with an anisotropic term based off the GGX Anisotropic model:
  \begin{lstlisting}
  const float Param  = sqr(0.5f + sqr(m->roughness)),
  Aspect = sqrt(1.0f - m->anisotropic*0.9f),
  Ax     = Param/Aspect, Ay = Param*Aspect,
  GGX_NV = Smith_G_GGX_Correlated(cos_HL, cos_NV, Ax),
  GGX_HL = Smith_G_GGX_Correlated(cos_NV, cos_HL, Ay);
  microfacet *= 0.5f / (GGX_NV*Ax + GGX_HL*Ay);
  \end{lstlisting}
  \includegraphics[scale=0.20]{geometric}

  And finally, the Distribution term controls how microfacets are spread around the normal. It's based off the Hyper Cauchy with a roughness and metallic (shape) parameters [9]. The latter of which should most likely be its own parameter in the future. The more readable mathematical model is shown:

  \begin{align}
  \frac{(M - 1)\:(\sqrt{2})^{2\:S\:-\:2}}
  	    {\pi\:R^{2}\:cos^{4}\theta\:(2 +
  	      \frac{tan^{2}\theta}{M^2})^M}
  \end{align}
  \includegraphics[scale=0.20]{distribution}

@@

  \section{Bibliography}
  [1]
  \begin{align*}
    \Sigma_{k=1}^{\infty} \int_{M^{k+1}} L_e(v_0 \rightarrow v_1) G(v_0
    \leftrightarrow v_1)
    \Pi_{i=1}^{k-1} F_s(v_{i-1} \rightarrow v_i \rightarrow v_{i+1}) G(v_i
    \leftrightarrow v_{i+1})\\
    W_e^j(v_{k-1} \rightarrow v_k) dA(v_0) \ldots dA(v_k)
  \end{align*}
  [9]
  Samuel D. Butler, Experimental and Theoretical Basis for a closed-Form Spectral BRDF model

>>>>

  Now, the importance of probability distribution functions comes when
performing Monte Carlo approximation on an integral. Take, for instance, the
most well-known form of the rendering equation:
  \begin{align}
    L_o(P, \omega_o) = L_e(P, \omega_o) + \int_{S}f_s(P, \omega_i,
    \omega_o) L_i(P, \omega_i) \vert cos\theta_i\vert d\omega_i
  \end{align}

  Where:
  \begin{list}{}{}
  	 \item $Lo(P,\:\omega_o)$ is the radiance at point $P$ in space towards direction $\omega_o$.

  	 \item $L_e(P,\:\omega_o)$ is the emittance at point $P$ towards $\omega_o$
      (such as a lamp).

  	 \item $S$ is the unit sphere in which the angle $\omega_i$ is integrated
        over in respects to the normal at $P$. Also, this means, implicitly,
        that the equation is in respects to solid angle; the angle at which a
        primitive $A$ subtends $B$.

  	 \item $f_s$ is the Bidirectional Scattering Distribution Function. This
        will be detailed in the BSDF chapter, suffice it to say, it describes
        how photons are scattered (or absorbed) from point P in space from
        direction $\omega_i$ towards $\omega_o$ -- either through reflection
        (BRDF), or transmittance (BTDF). In volumetric path tracing, there is
        also subsurface scattering, but this phenomena is only approximated in
        DTQ.

  	 \item $L_i(P, \omega_i)$ is the rendering equation at P from $\omega_i$

  	 \item And $cos\theta_i$ is the normal attenuation.
  \end{list}
  Conceptually, this is the behavior of photons bouncing off several surfaces of
a scene. While it's a very simple model, it's impossible to solve analytically
for anything but the most simple of scenes, mostly due to, for example, the
radiance at $P0$ relies on $P1$, and $P1$ relies on $P0$. Another thing to
remember is that while it is shown to be integrated over a sphere, it is usually
shown to be integrated over hemisphere $\Omega$ using either $f_r$ or $f_s$.

@@

  \begin{list}{}{}
  \item BSDF Spectrum: The albedo at $(P,\:\omega_i,\:\omega_o)$
  \item BSDF Sample: Samples a $\omega_o$ given $(P,\:\omega_i)$
  \item BSDF PDF: The probability distribution of $(P,\:\omega_i,\:\omega_o)$
  \end{list}

  There are many more optimizations to be made as well. For example, at each
scattering event, an additional photon can be sent towards each emitter. This is
what is known as next-event estimation. While this improves performance
drastically, naively averaging each contribution is incorrect, because each path
has a different probability to their paths/contributions. The technique to
sample them properly, multiple importance sampling, is related to BDPT, however
it will be detailed . The important thing to understand is that, in this
sampling strategy, there is a new PDF introduced which relates to the
probability of a path being generated, and this is critical to understand the
concept of BDPT.

  Now, the optimization that is of focus in DTQ is Bidirectional Path Tracing.
Instead of starting from an eye or light, a 'photon' (path) is generated from
both, allowed to scatter in the scene for a few paths (the length is determined
by Russian Roulette), and connected. It's not exactly the end-points of the
paths that are being connected, but each vertex generated is connected to each
other. [SHOW EXAMPLE]. This actually has an interesting correlation to
next-event estimation, as it could be seen as 'extending' the amount of emitters
that can be connected to.

  For BDPT, like a lot of different light transport equations, things go much
more smoothly when the equation uses a different measure. In this case, the
measure is over surface area rather than solid angles. This looks like $v_0
\rightarrow v_1$. The transformation looks like so:
  \begin{align}
    f_s(v \rightarrow \dot{v} \rightarrow \ddot{v}) &= f_s(P, \omega_i,
                                                      \omega_o)\\
    L_e(v \rightarrow \dot{v}) &= L_e(P, \omega_o)\\
    L_e(v \rightarrow \dot{v}) &= L_e^0(v) L_e^1(v \rightarrow \dot{v})
  \end{align}
    That is, instead of looking at a point $P$ with an incoming and outgoing
    angle $(\omega_i,\:\omega_o)$, look at the path from $P_{-1}$ ($v$) to
    $P_{+1}$ ($\ddot{v}$). More accurately, it's the transport between
    three different surfaces of the scene. There's a small change to be made in
the case of $L_e$, as DTOADQ [at least, for now], only has support for
area-light. Thus consider:

@@

    The $cos\theta_i$ is no longer applicable, instead there is now a geometry
term, G, that converts a PDF with respect to solid angle to surface area, which
involves the inverse squared distance, and the cosine angle between the
geometric normals at $v$ and $\dot{v}$:

@@

    Where $V$ is the visibility test. Thus the rendering equation with respects
    to area is:
  \begin{align}
    L(v \rightarrow \dot{v}) = L_e(v \rightarrow \dot{v}) \int_ML(v \rightarrow
    \dot{v}) G(v\leftrightarrow \dot{v}) f_s(v\rightarrow \dot{v} \rightarrow
    \ddot{v}) d_A
  \end{align}
    where $A$ is the area-measurement function on $M$, and $M$ is the set of all
points on the surface of a scene. That is, to say in regards to SDFs, $M$ is the
set of all points $P$ where $|map(P)| < \epsilon$. The area-measurement function
goes into measure theory, which while it is beyond my grasp currently to
compute, conceptually I understand a bit of it. The important thing is that $A$
is a function on the set of all subsets of $M$ that returns the measurement of
each set, and as long as the surfaces of the scene are sampled uniformly, this
never actually has to be computed. To back it up a little further, Veach
presented this as a means to integrate over a measure rather than using solid
angles. This proves that you can transform the rendering equation to different
measurements. For example, instead of integrating over points on surfaces, one
could integrate over an entire 3D space in order to perform volumetric path
tracing. Transforming the rendering equation using measurements would be
interesting in the future to look into.

    Anyways, with the three-point form now described, by recursively expanding,
so that the three point form is integrated over every possible set of possible
paths, this can be rewritten as the more usable format:
  \begin{align} I_j = \int_\Omega f_j(\bar{v}) d\mu(\bar{v}) \end{align}
    Where $\bar{v}$ is a path $z_0\rightarrow z_s \rightarrow y_t \rightarrow
    y_0 (eye \rightarrow light)$, $\Omega$ is the combination of all possible
    paths of any length, and $\mu$ is the area-product measure [the product of
all the expanded $A$]. Special care has to be taken with the definition of
    path itself, as in the source code, there never exists a path $\bar{v}$,
    only eye-path $\bar{z}$ and light-path . Monte Carlo approximation, as shown
above, is applied to the formula:
  \begin{align}
    I_j \approx \frac{1.0f}{N}\Sigma_0^NF\\
    F = \Sigma_s\Sigma_t\frac{f_j(\bar{v}_{s, t})}{p(\bar{v}_{s, t})}
  \end{align}

@@

    And now
  \begin{align}
    F = \Sigma_s\Sigma_t \alpha^L_s c_{s, t} \alpha^E_t
  \end{align}

  $\alpha_{s|t}^{L|E}$ is the final radiance of either the emitter or eye path,
  and thus it is recursive in nature. That is, the value of $\alpha_i$ is
  dependent on $\alpha_{i-1}$, unless $i = 0$ in which case is the radiance of
  either the emitter or the eye path. However, this can be simplified further,
because of Helmhotz reciprocity, BRDFs return the same value if $wo$ and $wi$
are switched. This is also the reason why forward and backward path tracing
work.
  \begin{align}
    \alpha^{L|E}_i =
    \begin{cases}
      1.0 &i = 0,\\
      \frac{
      L_e^0(y_0 \rightarrow y_1)}{p_A(y_1)} | 1.0 &i = 1,\\
      \frac{f_s(v_{i-1} \rightarrow v_i \rightarrow v_{i+1})}{p_{\sigma}(v_i
      \rightarrow v_{i-1})} &i > 1
    \end{cases}
  \end{align}
    For the case of $i = 0$, this would mean that the path does not exist. In
other words, forward or backward path tracing, the latter of which isn't
supported as camera lens are not defined in DTQ. For $i = 1$, it is either a
point on the surface of a light, or the origin point on the 'delta eye'. Lastly,
the BRDF is used to calculate the albedo of each primitive in the scene, and is
then divided by the PDF sigma, which is the probability density of $v_i
\rightarrow v_{i-1}$. $\alpha_1^L$ is shown below:

\begin{lstlisting}
  float3 N      = Sample_Cosine_Sphere(si, &pdf_pos),
         origin = light.origin + light.radius*N,
         dir    = Reorient_Angle(Sample_Cosine_Hemisphere(si, &pdf_dir), N);
  float pdf_pos = 1.0f/(2.0f*TAU*SQR(light.radius)),
        pdf_amt = 1.0f/EMITTER_AMT;
  pv->irradiance = light.emission/(pdf_pos*pdf_amt*pdf_fwd);
\end{lstlisting}

  Note that the irradiance is being calculated in this case, that is, the amount
  of flux arriving at point $pv$. Though in the case of $s=1$, this is more like
  the radiance. The irradiance at $t > 1$ and $s > 1$ is shown:

\begin{lstlisting}
  float3 sample_dir, sample_f;
  float sample_pdf;
  _BSDF_Sample(wi, N, vtx, &sample_dir, &sample_pdf, &sample_f, si);
  *irradiance = sample_f * (*irradiance) * fabs(dot(N, sample_dir))/sample_pdf;
\end{lstlisting}

  The connection strategy below describes how to connect the edges of the two
paths
  \begin{align}
    c_{s, t} =
    \begin{cases}
      L_e(z_{-2} \rightarrow z_{-1}), &s = 0, t > 1,\\
      f_s(y_{-2} \rightarrow y_{-1} \rightarrow z_{-1}) G(y_{-1} \leftrightarrow
      z_{-1}) f_s(y_{-1} \rightarrow z_{-1} \rightarrow z_{-2}), &s > 0, t > 1
    \end{cases}
  \end{align}
    Is the $s = 0, t > 1$ strategy worthwhile to compute? Those familiar with
next event estimation would know that there are situations in which this is
beneficial. Take, for example, a glossy surface with a small solid angle
subtending to the surface of a large light source. The $s=1$ sample strategy in
this case would only contribute when the $L0$ vertex is within that solid angle.
The effective strategy to sample in this case is $s = 0$. In code,
$\alpha_i^E\:c_{s,t}\:\alpha_0^L$ looks like (MIS not shown):
    \begin{lstlisting}
     float3 wi = normalize(E1.origin - E0.origin),
            wo = normalize(L1.origin - E1.origin);
     float3 albedo = _BRDF_F(wi, E1.normal, wo, E1.material, E1.mat_col);
     albedo /= BSDF_PDF(wo, wi, &E1);
     Spectrum contribution = E1.radiance * albedo;
     contribution *= Geometric_Term(&E1, &L1, si, Tx);
     contribution *= tlight.emission/Light_PDF(&tlight);
    \end{lstlisting}

@@

  The Geometric term is used to control how microfacets 'shadow' each other. The
DTQ BRDF uses Heits' [2014] SmithGGXCorrelated modified to use a half vector,
combined with an anisotropic term based off the GGX Anisotropic model:
  \begin{lstlisting}
  const float Param  = sqr(0.5f + sqr(m->roughness)),
  Aspect = sqrt(1.0f - m->anisotropic*0.9f),
  Ax     = Param/Aspect, Ay = Param*Aspect,
  GGX_NV = Smith_G_GGX_Correlated(cos_HL, cos_NV, Ax),
  GGX_HL = Smith_G_GGX_Correlated(cos_NV, cos_HL, Ay);
  microfacet *= 0.5f / (GGX_NV*Ax + GGX_HL*Ay);
  \end{lstlisting}
  \includegraphics[scale=0.20]{geometric}
  And finally, the Distribution term controls how microfacets are spread around
the normal. It's based off the Hyper Cauchy with a roughness and metallic
(shape) parameters [9]. The latter of which should most likely be its own
parameter in the future. The more readable mathematical model is shown:
  \begin{align}
  \frac{(M - 1)\:(\sqrt{2})^{2\:S\:-\:2}}
  	    {\pi\:R^{2}\:cos^{4}\theta\:(2 +
  	      \frac{tan^{2}\theta}{M^2})^M}
  \end{align}
  \includegraphics[scale=0.20]{distribution}

@@

  \section{Bibliography}
  [9]
  Samuel D. Butler, Experimental and Theoretical Basis for a closed-Form Spectral BRDF model
